---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# collatRal

<!-- badges: start -->
<!-- badges: end -->

The goal of collatRal is to test collateral effects in clinical surveillance data, with antibiotic MIC measurements for a set of strains. The package currently includes a main test functions, a wrapper for performing the test for multiple combinations of antibiotics and two plot functions.

## Installation

The package can be installed directly from github using `devtools`. Run the following line in R:
```{r installation, echo = TRUE, eval = FALSE}
devtools::install_github("vanhasseltlab/collatRal")
```


## Example

A typical data set includes rows with strains and columns with drugs/antibiotic

```{r example}
library(collatRal)
MIC_test <- data.frame(d1 = c(16, 16, 128, 128, 32, 16, 128, 1, 64, 1), 
                       d2 = c(32, 32, 16, 32, 8, 32, 8, 8, 8, 8), 
                       d3 = c(8, 64, 64, 64, 16, 6, 16, 8, 8, 8))
```

You can test collateral effects between different drugs using the core function `collateral_t_test()`.

```{r}
MIC_test_result_12 <- collateral_t_test(MIC_test$d1, MIC_test$d2)
print(MIC_test_result_12)
```

To visualize this result you can use the plot the histogram:

```{r}
plot_histogram_CE(MIC_test$d1, MIC_test$d2)
```

In order to test more collateral drug effects, you can use the wrapper `collateral_mult_test`, which will give the results over different antibiotic combinations.

```{r}
MIC_test_result_all <- collateral_mult_test(MIC_test)
```

To visualize these results we have implemented a heatmap plot as well.

```{r}
plot_heatmap_CE(MIC_test_result_all)
```

Please check the manual for more options for the test.
